### 2018-04-02 equal_example.tex
initial commit

\begin{tabular}{C{0.66in}C{0.66in}C{0.66in}C{0.66in}C{0.66in}C{0.66in}C{0.66in}}
\toprule
 experiment &  human droplets &  mouse droplets &  nonempty droplets &  human and mouse droplets &  multiplet freq &  twice cross celltype freq \\
\midrule
          1 &            2005 &            2005 &               4000 &                        10 &           0.005 &                      0.005 \\
          2 &            2050 &            2050 &               4000 &                       100 &           0.049 &                      0.050 \\
          3 &            2500 &            2500 &               4000 &                      1000 &           0.425 &                      0.500 \\
\bottomrule
\end{tabular}